A general Nullstellensatz for generalized spaces

### 2019-03-17 paper-qcoh.tex
Terminology clarification

A fragment of logic which is still finitary, but is powerful enough to express
its own provability predicate, is coherent existential fixed-point
logic~\cite{blass:topoi-and-computation,blass:existential-fixed-point-logic}.
For this fragment, we have the following characterization.

\begin{scholium}Let~$\TT$ be a coherent theory (or more generally theory in coherent
existential fixed-point logic). Let~$\alpha$ be a finitary first-order
formula over the signature of~$\TT$. Then the following statements are
equivalent.
\begin{enumerate}
\item The formula~$\alpha$ holds for~$U_\TT$. \smallskip
\item The formula~$\alpha$ is provable in coherent existential fixed-point logic
modulo the axioms of~$\TT$ and the additional axioms
\begin{equation}
  \speak{$\sigma$ holds for~$U_\TT$} \Longrightarrow \speak{$\ul{\TT}/U_\TT$
  proves~$\sigma$}
\end{equation}
where~$\sigma$ ranges over the formulas of coherent existential fixed-point
logic over the signature of~$\TT$.
\end{enumerate}
\end{scholium}

>>>>

A fragment of logic which is still finitary, but is powerful enough to express
its own provability predicate, is \emph{coherent existential fixed-point
logic}, coherent logic enriched by list sorts and the fixed-point operator of
existential fixed-point logic~\cite{blass:topoi-and-computation,blass:existential-fixed-point-logic}.
The list sorts are necessary to express raw terms, raw formulas and raw
sequents; the fixed-point operator is then used to express well-formedness of
raw terms, raw formulas and raw sequents; and to express provability. For this
fragment, we have the following characterization.

\begin{scholium}Let~$\TT$ be a coherent theory (or more generally theory in coherent
existential fixed-point logic). Let~$\alpha$ be a finitary first-order
formula over the signature of~$\TT$. Then the following statements are
equivalent.
\begin{enumerate}
\item The formula~$\alpha$ holds for~$U_\TT$. \smallskip
\item The formula~$\alpha$ is provable in coherent existential fixed-point logic
modulo the axioms of~$\TT$ and the additional axioms
\[
  \speak{$\sigma$ holds for~$U_\TT$} \Longrightarrow \speak{$\ul{\TT}/U_\TT$
  proves~$\sigma$}
\]
where~$\sigma$ ranges over the formulas of coherent existential fixed-point
logic over the signature of~$\TT$.
\end{enumerate}
\end{scholium}
